Kaluza--Klein Cosmology from five-dimensional Lovelock--Cartan Theory

### 2016-09-14 KKosmology.tex
Some modifications of the plots and their calls

Due to Eq.~\eqref{u}, the solutions will be valid in the region of the parameter space where $\Delta>0$. The function $\omega(t)$ satisfies the equation
\begin{align}
  \dot{\omega}+\omega^2+u_{\pm} &= 0,
\end{align}
which, for the three significantly different values of $u_\pm$, has the following solutions
\begin{align}
  u_\pm>0&:\omega(t)=-\sqrt{u_\pm}\tan\left[\sqrt{u_\pm}\left(t-t_0\right)\right],\\
  u_\pm=0&:\omega(t)=\left(t-t_0\right)^{-1},\\
  u_\pm<0&:\omega(t)=\sqrt{-u_\pm}\tanh\left[\sqrt{-u_\pm}\left(t-t_0\right)\right],
\end{align}
where $t_0$ is an integration constant to be fixed. We express the time-dependence of the remaining fields in terms of $\omega$ and list them explicitly in Appendix~\ref{solutions t}. The solutions read
\begin{align}
  \gamma(\omega)&=\frac{u_\pm}{\omega},\\
  \phi(\omega)&=\frac{\phi_0\,\omega^2}{\omega^2+u_\pm}\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2\left(\omega^2+u_\pm\right)}\right],\\
  a(\omega)&=\frac{a_0}{\sqrt{\omega^2+u_\pm}}\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2\left(\omega^2+u_\pm\right)}\right],\\
  f^2(\omega)&=\left(\omega^2+u_\pm\right)\frac{k}{a_0^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2\left(\omega^2+u_\pm\right)}\right]\notag\\
  &+\omega^2-Y_\pm,
\end{align}
where we have defined the constant
\begin{equation*}
  Y_{\pm} = -\frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2},
\end{equation*}
and $a_0$ and $\phi_0$ are integration constants.

@@

For $u_\pm>0$ (see Fig.~\ref{uposreg}) the solutions are periodic. In that case, the scale factor $a(t)$ starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$. Depending on the particular region on the parameter space, the scale factor undergoes an expanding and contracting age, allowing to reach a minimum value without collapsing, to expand again before a big crunch. Otherwise, it reaches a maximum value before its collapse. It also shows an inflationary phase at early times when the scale factor grows exponentially. The bouncing behaviour of the extra dimension is given by the scalar field in a phase difference of $\pi/2$ with respect to $a(t)$.

\begin{figure}[H]
  \includegraphics[scale=.45]{cond_u_pos.pdf}
  \caption{The figure on the left denotes the region on the parameter space where $u_+>0$, while the figure on the right shows where $u_->0$. We have normalized $\alpha_1 = 1$.}
  \label{uposreg}
\end{figure}

\begin{widetext}
  \begin{figure}[H]
    \includegraphics[scale=.45]{a_u_pos.pdf} \includegraphics[scale=.45]{phi_u_pos.pdf}
    \caption{blablabla}
    \label{upossol}
  \end{figure}

\end{widetext}

% \begin{figure}[H]
%  \includegraphics[scale=.45]{a_u_pos.pdf}
% \label{aupos}
% \end{figure}
% 
% \begin{figure}[H] 
% \includegraphics[scale=.45]{phi_u_pos.pdf}
%  \caption{These plots show the bouncing behaviour of the solutions in the distinctive regions of the parameter space. We have normalized each plot to its maximum value $a_{\rm max}$ and $\phi_{\rm max}$, respectively.}
%  \label{phiupos}
% \end{figure}

@@

For $u_\pm\leq 0$ (see Fig.~\ref{unegreg}), depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reaches an asymtotic vanishing value for late cosmic times, after a period of expansion. This is interpreted as the late time behaviour of the radius of the compact manifold $S^1$, which for this family of solutions provides a natural explanation for its scale. Otherwise, in contrast to the previous behaviour, the scale factor contracts to an asymtotic vanishing value while the scalar field has no upper bound. In the case of strictly negative $u_\pm$, for either contracting or expanding scale factors, their initial values remain finite and the solutions do not present an initial singularity. It has been reported that the Gauss--Bonnet term can prevent the Universe to expand from an initial singularity \cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also the present case.

\begin{figure}[H]
  \includegraphics[scale=.45]{cond_u_neg.pdf}
  \caption{The figure on the left denotes the region on the parameter space where $u_+<0$, while the figure on the right shows where $u_-<0$. We have normalized $\alpha_1 = 1$.}
  \label{unegreg}
  \includegraphics[scale=.45]{a_u_neg.pdf} 
  \includegraphics[scale=.45]{phi_u_neg.pdf}
\end{figure}

% \begin{figure}[H]
%  \includegraphics[scale=.45]{a_u_neg.pdf} 
%  \includegraphics[scale=.45]{phi_u_neg.pdf}
%  \caption{Behaviour for $u_-<0$.}
%  \label{unegsol}
% \end{figure}

\begin{figure}[H]
  \includegraphics[scale=.45]{a_u-_0.pdf} 
  \includegraphics[scale=.45]{phi_u-_0.pdf}
  \caption{Behaviour for $u_- = 0$.}
  \label{uzerosol}
\end{figure}

@@

$\bullet\ u_\pm>0$:
\begin{align}
  \gamma(t)&=-\sqrt{u_\pm}\cot\tau,\\
  \phi(t)&=\phi_0\sin^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2 u_\pm}\cos^2\tau\right],\\
  a(t)&=\frac{a_0}{\sqrt{u_\pm}}|\cos\tau|\exp\left[\frac{\pm\sqrt{6\Delta}}{24\alpha_2 u_\pm}\cos^2\tau\right],\\
  f^2(t)&=u_\pm\frac{k}{a_0^2}\sec^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cos^2\tau\right]\notag\\
  &+u_\pm\tan^2\tau-Y_\pm.
\end{align}

$\bullet\ u_\pm=0$:
\begin{align}
  \gamma(t)&=0,\\
  \phi(t)&=\phi_0\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2}(t-t_0)^2\right],\\
  a(t)&=a_0|t-t_0|\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2}(t-t_0)^2\right],\\
  f^2(t)&=\frac{k}{a_0^2(t-t_0)^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2}(t-t_0)^2\right]\notag\\
  &+(t-t_0)^{-2}-Y_\pm.
\end{align}

$\bullet\ u_\pm<0$:
\begin{align}
  \gamma(t)&=-\sqrt{-u_\pm}\coth\tau,\\
  \phi(t)&=\phi_0\sinh^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2u_\pm}\cosh^2\tau\right],\\
  a(t)&=\frac{a_0}{\sqrt{-u_\pm}}\cosh\tau\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2u_\pm}\cosh^2\tau\right],\\
  f^2(t)&=-u_\pm\frac{k}{a_0^2}\cosh^{-2}\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cosh^2\tau\right]\notag\\
  &-u_\pm\tanh^2\tau-Y_\pm.
\end{align}
%%%%%%%%% BIBLIOGRAPHY %%%%%%%%%
%%\bibliographystyle{apsrev4-1}
\bibliography{References.bib}

>>>>

Due to Eq.~\eqref{u}, the solutions will be valid in the region of the parameter space where $\Delta>0$. The function $\omega(t)$ satisfies the equation
\begin{align}
  \dot{\omega}+\omega^2+u_{\pm} &= 0,
\end{align}
which, for the three significantly different values of $u_\pm$, has the following solutions
\begin{align}
  u_\pm>0&:\omega(t)=-\sqrt{u_\pm}\tan\left[\sqrt{u_\pm}\left(t-t_0\right)\right],\\
  u_\pm=0&:\omega(t)=\left(t-t_0\right)^{-1},\\
  u_\pm<0&:\omega(t)=\sqrt{-u_\pm}\tanh\left[\sqrt{-u_\pm}\left(t-t_0\right)\right],
\end{align}
where $t_0$ is an integration constant to be fixed. We express the time-dependence of the remaining fields in terms of $\omega$ and list them explicitly in Appendix~\ref{solutions t}. The solutions read
\begin{align}
  \gamma(\omega)&=\frac{u_\pm}{\omega},\\
  \phi(\omega)&=\frac{\phi_0\,\omega^2}{\omega^2+u_\pm}\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2\left(\omega^2+u_\pm\right)}\right],\\
  a(\omega)&=\frac{a_0}{\sqrt{\omega^2+u_\pm}}\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2\left(\omega^2+u_\pm\right)}\right],\\
  f^2(\omega)&=\left(\omega^2+u_\pm\right)\frac{k}{a_0^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2\left(\omega^2+u_\pm\right)}\right]\notag\\
  &+\omega^2  + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2},
\end{align}
where $a_0$ and $\phi_0$ are integration constants.

@@

For $u_\pm>0$ (see Fig.~\ref{cond_u_pos}) the solutions are periodic. In that case, the scale factor $a(t)$ starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$. Depending on the particular region on the parameter space, the scale factor undergoes an expanding and contracting age, allowing to reach a minimum value without collapsing, to expand again before a big crunch. Otherwise, it reaches a maximum value before its collapse. It also shows an inflationary phase at early times when the scale factor grows exponentially. The bouncing behaviour of the extra dimension is given by the scalar field in a phase difference of $\pi/2$ with respect to $a(t)$.

\begin{figure}[H]
  \includegraphics[width=\linewidth]{cond_u_pos.pdf}
  \caption{The figure on the left denotes the region on the parameter space where $u_+>0$, while the figure on the right shows where $u_->0$. We have normalized $\alpha_1 = 1$.}
  \label{cond_u_pos}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=\linewidth]{a_u_pos.pdf}
  \caption{blablabla}
  \label{a_u_pos}
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{phi_u_pos.pdf}
  \caption{Behavior of the $\phi$ field as a function of the scaled time, for $u_\pm > 0$.}
  \label{phi_u_pos}
\end{figure}

@@

For $u_\pm\leq 0$ (see Fig.~\ref{cond_u_neg}), depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reaches an asymtotic vanishing value for late cosmic times, after a period of expansion. This is interpreted as the late time behaviour of the radius of the compact manifold $S^1$, which for this family of solutions provides a natural explanation for its scale. Otherwise, in contrast to the previous behaviour, the scale factor contracts to an asymtotic vanishing value while the scalar field has no upper bound. In the case of strictly negative $u_\pm$, for either contracting or expanding scale factors, their initial values remain finite and the solutions do not present an initial singularity. It has been reported that the Gauss--Bonnet term can prevent the Universe to expand from an initial singularity \cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also the present case.

\begin{figure}[H]
  \includegraphics[scale=.45]{cond_u_neg.pdf}
  \caption{The figure on the left denotes the region on the parameter space where $u_+<0$, while the figure on the right shows where $u_-<0$. We have normalized $\alpha_1 = 1$.}
  \label{cond_u_neg}  
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{a_u_neg.pdf}
  \caption{}
  \label{a_u_neg}
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{phi_u_neg.pdf}
  \caption{}
  \label{phi_u_neg}
\end{figure}

\begin{figure}[H]
  \includegraphics[scale=.45]{a_u-_0.pdf} 
  \caption{Behaviour for $u_- = 0$.}
  \label{a_u_0}
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{phi_u-_0.pdf}
  \caption{Behaviour for $u_- = 0$.}
  \label{phi_u_0}
\end{figure}

@@

$\bullet\ u_\pm>0$:
\begin{align}
  \gamma(t)&=-\sqrt{u_\pm}\cot\tau,\\
  \phi(t)&=\phi_0\sin^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2 u_\pm}\cos^2\tau\right],\\
  a(t)&=\frac{a_0}{\sqrt{u_\pm}}|\cos\tau|\exp\left[\frac{\pm\sqrt{6\Delta}}{24\alpha_2 u_\pm}\cos^2\tau\right],\\
  f^2(t)&=u_\pm\frac{k}{a_0^2}\sec^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cos^2\tau\right]\notag\\
  &+u_\pm\tan^2\tau - \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
\end{align}

$\bullet\ u_\pm=0$:
\begin{align}
  \gamma(t)&=0,\\
  \phi(t)&=\phi_0\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2}(t-t_0)^2\right],\\
  a(t)&=a_0|t-t_0|\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2}(t-t_0)^2\right],\\
  f^2(t)&=\frac{k}{a_0^2(t-t_0)^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2}(t-t_0)^2\right]\notag\\
  &+(t-t_0)^{-2} + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
\end{align}

$\bullet\ u_\pm<0$:
\begin{align}
  \gamma(t)&=-\sqrt{-u_\pm}\coth\tau,\\
  \phi(t)&=\phi_0\sinh^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2u_\pm}\cosh^2\tau\right],\\
  a(t)&=\frac{a_0}{\sqrt{-u_\pm}}\cosh\tau\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2u_\pm}\cosh^2\tau\right],\\
  f^2(t)&=-u_\pm\frac{k}{a_0^2}\cosh^{-2}\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cosh^2\tau\right]\notag\\
  &-u_\pm\tanh^2\tau + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
\end{align}
%%%%%%%%% BIBLIOGRAPHY %%%%%%%%%
%%\bibliographystyle{apsrev4-1}
\bibliography{References.bib}
